Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2014-12-16 doc/GMP2015/s02_preliminaries.tex
Added a Basics section (incomplete)

\subsection{Definitions}

% TODO

% A paragraph of preliminaries such as:
%     - rigidity matrix R,
%     - stress vectors S one coordinate per edge (w/ SR =E (external stress vector, 2 coordinates  per vertex));
%     - flex vectors F (2 coordinates per vertex) (right nullspace/kernel of R)
% Should precede the definition of
%     + independent,
%     + overconstrained (not independent),
%     - rigid (contains an independent set with sufficiently many eedges/constraints)
%     - isostatic (wellconstrained/minimally rigid), ,
%     - flexible (not rigid),
%     + underconstrained (independent and not rigid).
% After that, define DR plans etc.

In Sections \ref{sec:DRP} and \ref{sec:recomb} a \textbf{graph} is a bar and joint system. $G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$ (undirected). Additionally, there are two weight functions, $w_V: V \to \mathbb{R}^+$ and $w_E: E \to \mathbb{R}^+$. The \textbf{density} of graph $G$ is $d(G) = \sum_{e\in E}{w_E(e)} - \sum_{v\in V}{w_V(v)}$.

Given a constant $k$, a graph $G$ is:
\begin{itemize}
    \item \textbf{independent} (or \textbf{sparse}) if, for all non-trivial subgraphs $S\subseteq G$, $d(S) \leq k$.

    \item \textbf{over-constrained} if it is not independent.

    \item \textbf{well-constrained} (or \textbf{tight}) if $G$ is independent and $d(G)=k$.

    \item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.

    \item \textbf{under-constrained} if $G$ is independent and not rigid.

    % \item \textbf{trivial} if (1) it is over-constrained graph and (2) all of its subgraphs are also trivial. Trivial graphs are input.
\end{itemize}

\textbf{Trivial} graphs are ill-defined in the general case. The only strict requirements are: (1) it must over-constrained and (2) all of its subgraphs are also trivial.

>>>>

\subsection{Basics}
%TODO!

%Standard except for generality and ordering and the fact that I want to
%avoid the use of the word framework if possible. Do everything
%with constraint systems, linearization and genericity using indeterminates.

%--general  geometric constraint system --

A $d$-dimensional {\it geometric constraint system} in general consists of

- a underlying {\it constraint graph} $G=(V,E)$, or more generally, a hypergraph, whose  vertices represents geometric primitives in $d$-dimension and whose (hyper)edges specifies the combinatorics of constraints between pairs/tuples of the geometric primitives.

- an {\it algebraic system of  constraint functions} from $E$ to the domain of constraint values

In this paper we study qusecs, a class of 2D geometric constrain system that is independent (defined below).

A familiar example of geometric constraint system is {\it linkage (barjoint system)},
where the underlying constraint graph is a simple graph and
the constraints are Euclidean distances on the edges of the graph.

%--Give example as
%linkage(barjoint system) 
\todo{-- refer to others (citations
+ pinned line incidence and multi-body-pin (in application
section)}

%--underlying constraint graph or hypergraph (common treatment)
\todo{???The (hyper)graph of a is sometimes refered to???}

%--linearization: Jacobian of constraint system,
%rigidity matrix of indeterminates

The algebraic system can usually be linearized at {\it generic} points
by taking the Jacobian of the system,
giving  the $d$-dimensional {\it rigidity matrix } $R_d(G)$ of a the (hyper)graph $G = (V,E)$, 
whose rows correspond to edges, columns correspond to coordinate positions of vertices,  entries are \emph{indeterminates} representing the coordinate position $\mathbf{p}(v) \in \mathbb{R}^d$ of the geometric primitives corresponding to vertices $v \in V$.

%-- give example for the case of
%linkages

In the case of linkage, 
let $p_1(v), p_2(v), \ldots, p_d(v)$ represent the coordinate position $\mathbf{p}(v) \in \mathbb{R}^d$ of the joint corresponding to a vertex $v \in V$. 
The matrix $R_d(G)$ has one row for each edge $e \in E$ and $d$ columns for each vertex $v \in V$.  The row corresponding to $e = \{u,v\} \in E$ represents the bar connecting $p(u)$ and $p(v)$ and has $d$ non-zero indeterminate entries $\mathbf{p}(u)-\mathbf{p}(v)$ (resp.\ $\mathbf{p}(v)-\mathbf{p}(u)$) in the $d$ columns corresponding to $u$ (resp.  $v$) and zero in the other entries.

%%Let $p_1(v), p_2(v), \ldots, p_d(v)$ represent the coordinate position $\mathbf{p}(v) \in \mathbb{R}^d$ of the joint corresponding to a vertex $v \in V$. 
%The matrix $R_d(G)$ has one row for each edge $e \in E$ and $d$ columns for each vertex $v \in V$.  The row corresponding to the hyperedge $e = \{v_1, v_2, \ldots, v_k\} \in E$ represents the constraint between the vertices $ \{v_1, v_2, \ldots, v_k\}$, 
%has $d$ non-zero indeterminate entries %$\mathbf{p}(u)-\mathbf{p}(v)$ (resp.\ $\mathbf{p}(v)-\mathbf{p}(u)$) 
%in the $d$ columns corresponding to each vertex $v_i \in e$ and zero in the other entries.

%generic independence, infinitesimal rigidity
%(based on indeterminates)

A subset of edges $E'$, or a subgraph $(V', E')$, of a graph $G = (V, E)$ is said to be {\it independent} in $d$-dimensions, when the set of rows of $R_d(G)$ corresponding to $E'$ is generically independent, or independent for a generic instantiation of the indeterminate entries. 
This yields the $d$-dimensional generic {\it rigidity matroid} associated with a graph $G$.

The graph is  {\it infinitesimally rigid} if %the rigidity matrix has full rank, that is, 
if the number of generically independent rows or the rank of $R_d(G)$ is maximal.
In the case of linkage, the maximal rank is $d|V| - {d+1 \choose 2}$, where ${d+1 \choose 2}$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^d$ [?].
A vector in the kernel of the rigidity matrix is called an {\it infinitesimal motion vector}.

%--stress vectors and flex vectors
%(also based on indeterminates)

The left nullspace of the rigidity matrix is called the {\it space of stress}, consisting of the {\it flex vectors}.

%--A generalized version of Asimow Roth (implicit fn thm)
%generic rigidity equivalent to infinitesimal rigidity (easy
%when the latter def is based on indeterminates)

 A generalized version of Asimow Roth (implicit fn thm) states that
 generic rigidity, i.e.\ there exist at most finitely many solutions to the system at generic points,  %to the algebraic system 
  is equivalent to  our definition  of infinitesimal rigidity based on indeterminates.

%--generic isostaticity (minimal rigidity, wellconstrainedness),
%flexibility (not rigid), underconstrained (independent
%and not rigid), overconstrained (dependent -- only briefly
%mentioned
%in this paper; all results are about independent - underconstrained or
%isostatic)

A graph $G = (V,E)$ is:
\begin{itemize}
	\item generically {\it isostatic (minimally rigid, well-constrained)} if $R_d(G)$ has full rank
	\item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
	\item {\it flexible} if $G$ is not rigid
    \item \textbf{under-constrained} if $G$ is independent and not rigid.
    \item \textbf{over-constrained} if it is not independent.
\end{itemize}
In this paper all results are about independent - underconstrained or
isostatic

\todo{--rigidity matroid\\
--Combinatorial characterization of rigidity matroid
Example: Laman's theorem for bar-joint (refer to others)
Lovasz Yemini
Tay white and whiteley
(citations)
+ pinned line incidence and multi-body-pin (in application
section)}

% % % % % % % % % % % % % % % % % % % % % % % % % % % % %

%The problem is equivalent to combinatorially determining the generic rank of the 3-dimensional bar-joint rigidity matrix of a graph.  The $d$-dimensional bar-joint rigidity matrix of a graph $G = (V,E)$, denoted $R_d(G)$, is a matrix of indeterminates. Let $p_1(v), p_2(v), \ldots, p_d(v)$  represent the coordinate position $\mathbf{p}(v) \in \mathbb{R}^d$of the joint corresponding to a vertex $v \in V$.  . The matrix $R_d(G)$ has one row for each edge $e \in E$ and $d$ columns for each vertex $v \in V$.  The row corresponding to $e = \{u,v\} \in E$ represents the bar connecting $p(u)$ and $p(v)$ and has $d$ non-zero indeterminate entries $\mathbf{p}(u)-\mathbf{p}(v)$ (resp.\ $\mathbf{p}(v)-\mathbf{p}(u)$) in the $d$ columns corresponding to $u$ (resp.  $v$) and zero in the other entries.  
%
%A subset of edges $E'$, or a subgraph $(V', E')$, of a graph $G = (V, E)$ is said to be independent (we drop ``bar-joint'' from now on) in $d$-dimensions, when the set of rows of $R_d(G)$ corresponding to $E'$ is generically independent, or independent for a generic instantiation of the indeterminate entries. This yields the $d$-dimensional generic rigidity matroid associated with a graph $G$. The graph is rigid if the number of generically independent rows or the rank of $R_d(G)$ is maximal, i.e., $d|V| - {d+1 \choose 2}$, where ${d+1 \choose 2}$ is the number of rotational and translational degrees-of-freedom of a rigid body in $\mathbb{R}^d$ [?].

\subsection{DR-plan}
%\subsection{Definitions}
%\label{sec:prelim_defs}
%
%% TODO
%
%% A paragraph of preliminaries such as:
%%     - rigidity matrix R,
%%     - stress vectors S one coordinate per edge (w/ SR =E (external stress vector, 2 coordinates  per vertex));
%%     - flex vectors F (2 coordinates per vertex) (right nullspace/kernel of R)
%% Should precede the definition of
%%     + independent,
%%     + overconstrained (not independent),
%%     - rigid (contains an independent set with sufficiently many eedges/constraints)
%%     - isostatic (wellconstrained/minimally rigid), ,
%%     - flexible (not rigid),
%%     + underconstrained (independent and not rigid).
%% After that, define DR plans etc.
%
%In Sections \ref{sec:DRP} and \ref{sec:recomb} a \textbf{graph} is a bar and joint system. $G=(V,E,w_V,w_E)$ is a set of vertices, $V$, and a set of edges, $E$, defined as a tuple of vertices from $V$ (undirected). Additionally, there are two weight functions, $w_V: V \to \mathbb{R}^+$ and $w_E: E \to \mathbb{R}^+$. The \textbf{density} of graph $G$ is $d(G) = \sum_{e\in E}{w_E(e)} - \sum_{v\in V}{w_V(v)}$.
%
%Given a constant $k$, a graph $G$ is:
%\begin{itemize}
%    \item \textbf{independent} (or \textbf{sparse}) if, for all non-trivial subgraphs $S\subseteq G$, $d(S) \leq k$.
%
%    \item \textbf{over-constrained} if it is not independent.
%
%    \item \textbf{well-constrained} (or \textbf{tight}) if $G$ is independent and $d(G)=k$.
%
%    \item \textbf{rigid} if there exists some spanning subgraph $S\subseteq G$ such that $S$ is well-constrained.
%
%    \item \textbf{under-constrained} if $G$ is independent and not rigid.
%
%    % \item \textbf{trivial} if (1) it is over-constrained graph and (2) all of its subgraphs are also trivial. Trivial graphs are input.
%\end{itemize}
%
\textbf{Trivial} graphs are ill-defined in the general case. The only strict requirements are: (1) it must over-constrained and (2) all of its subgraphs are also trivial.

### 2014-12-16 doc/GMP2015/s08_pinnedline.tex
Added a Basics section (incomplete)

The optimal qusecs DR-plan can also be applied to analyze and solve
the structure of  cross-linking microfibrils in collagen and cellulose in biology,
by modeling these structures as {\em pinned line incidence systems}.

{\em Collagen} is an important protein material in biological tissues with highly elastic mechanical property.
{\em Cellulose} is the most important constituent of the cell wall of plants (see Figure~\ref{fig:cellulose}).
Both of these substances consist of a large number {\em microfibrils},
%fibril itself straight?
%Each fibril is attached to some fixed larger organelle/membrane at one spot (typically on the boundary).
%It is additionally
each of which is cross-linked at 2 places with some two other fibrils,
where the {\em  cross-linking} is like an incidence constraint that
the two crosslinked fibrils can slide with each other  \cite{xxx} (see Figure~\ref{fig:crosslink}).

@@

%TODO ???
The recombination problem for 2-dimensional pinned line incidence systems is trivial.
Since the pins are given fixed positions in the plane,
 the solutions of a well-constrained sub-system will automatically be consistent with solutions of the remaining of the system.

>>>>

The optimal qusecs DR-plan can also be applied to analyze and solve
the structure of  cross-linking microfibrils in collagen and cellulose in biology,
by modeling these structures as {\it pinned line incidence systems}.

{\it Collagen} is an important protein material in biological tissues with highly elastic mechanical property.
{\it Cellulose} is the most important constituent of the cell wall of plants (see Figure~\ref{fig:cellulose}).
Both of these substances consist of a large number {\it microfibrils},
%fibril itself straight?
%Each fibril is attached to some fixed larger organelle/membrane at one spot (typically on the boundary).
%It is additionally
each of which is cross-linked at 2 places with some two other fibrils,
where the {\it  cross-linking} is like an incidence constraint that
the two crosslinked fibrils can slide with each other  \cite{xxx} (see Figure~\ref{fig:crosslink}).

@@

\begin{remark}
The recombination problem for 2-dimensional pinned line incidence systems is trivial.
Since the pins are given fixed positions in the plane,
 the solutions of a well-constrained sub-system will automatically be consistent with solutions of the remaining of the system.
 \end{remark}
